---
title: "Capstone"
author: "Sahil Thapar"
date: "5/22/2017"
output: 
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = F, message = F, cache = T)
```

## The plan

1. Data cleaning
    + Identify missing values and outliers and handle them
2. Exploratory Data Analysis
    + Univariate analysis
        * Analyze the dependent variable, `SalesPrice`
    + Multivariate analysis
        * Analyze relationship between dependent and independent variables
            * numeric variables - correlation
            * categorical variables - distributrion across categories
3. Identify relatively important features based on a combination of instinct and numerical analysis.


## Data cleaning

```{r}
# packages
library(dplyr)
library(tidyr)
library(ggplot2)
library(tidyverse)
library(readr)
library(forcats)
library(purrr)
library(tibble)
library(corrplot)
library(scales)


# Read data
train <- read_csv('data/train.csv')

# Count missing values
missing.count <- colSums(sapply(train, is.na))
missing.count <- missing.count[missing.count > 0]
missing.count <- as_tibble(data.frame(column = names(missing.count),
                           count = missing.count,
                           row.names = NULL))

missing.count <-
  missing.count %>%
  mutate(percent_missing = round(count / nrow(missing.count))) %>%
  arrange(desc(count))
  
missing.count
```

* __PoolQC, MiscFeature, Alley, Fence, FireplaceQu__:  
    + All categorical variables where NA represents absence of the feature and not missing data.
* __Garage Variables__
    + Data is missing when no garage is present 
    * __GarageType, GarageFinish, GarageQual, GarageCond__
        * NA can be taken as a different category
    * __GarageYrBit__
        * We can remove this variable
* __Basement Variables__
    + Missing data represents no basement
    * __BsmtExposure, BsmtFinType1, BsmtFinType1, BsmtQual, BsmtCond__
        * NA can be taken as a different category
* __MasVnrType__
    + Masonry veneer type, NA represents no veneer
* __MasVnrArea__
    + Masonry veneer area, NA represents no veneer we can substitute with 0.
* __Electrical__
    + 1 missing observation. We can delete this observation
* __LotFrontage__
    + Can't set this to zero as it doesn't make sense that a house has no distance between the street
    
```{r}
create_none <- function(colName) {
  train[colName] <<- ifelse(is.na(train[colName]),
                           "None",
                           train[colName])
} 

# Remove id, garage year built, set masvnrarea to zero where null and remove observation with electrical as NA
train <- 
  train %>%
  select(-c(GarageYrBlt, Id)) %>%
  mutate(MasVnrArea = ifelse(is.na(MasVnrArea),
                                   0,
                                   MasVnrArea)) %>%
  filter(!is.na(Electrical)) 
```

To handle LotFrontage, it seems instinctive that it will be related to LotArea

```{r}
train %>%
  select(c(LotArea, LotFrontage)) %>%
  na.omit %>%
  cor

train %>%
  select(c(LotArea, LotFrontage)) %>%
  na.omit %>%
  mutate(LotArea = sqrt(LotArea)) %>%
  cor
```

We will substitue missing LotFrontage value with the square root of the lotarea

```{r}
train <-
  train %>%
  mutate(LotFrontage = ifelse(is.na(LotFrontage),
                              sqrt(LotArea),
                              LotFrontage))

train[is.na(train)] <- "None"
```

## Exploring the data

### Univariate analysis

```{r}
summary(train$SalePrice)
```


```{r}
  ggplot(data = train) +
  geom_histogram(mapping = aes(SalePrice),
                 fill = "#3498db",
                 color = "#2980b9",
                 binwidth = 10000) +
  scale_x_continuous(labels = comma_format()) +
  geom_vline(mapping = aes(xintercept = mean(train$SalePrice)),
                           colour = '#e74c3c',
                           size = 0.5) +
  geom_vline(mapping = aes(xintercept = median(train$SalePrice)),
                           colour = '#f39c12',
                           size = 0.5) +
  labs(x = "Sale Price",
       y = "Count")

  ggplot(data = train) +
  geom_histogram(mapping = aes(log(SalePrice)),
                 fill = "#3498db",
                 color = "#2980b9") +
  scale_x_continuous(labels = comma_format()) +
  geom_vline(mapping = aes(xintercept = log(mean(train$SalePrice))),
                           colour = '#e74c3c',
                           size = 0.5) +
  geom_vline(mapping = aes(xintercept = log(median(train$SalePrice))),
                           colour = '#f39c12',
                           size = 0.5) +
  labs(x = "Log Sale Price",
       y = "Count")
```

### Relationship with other numeric variables

```{r, fig.height = 12, fig.width = 10}
nums <- sapply(train, is.numeric)
M <- cor(train[, nums])
corrplot(M, method = "circle", type = "upper", order = "hclust")
```

For variable selection we can choose only the variables which have at least a correlation of 0.5 with SalesPrice

```{r}
M.tbl <- as_tibble(data.frame(variable = rownames(M),
                              correlation = M[, 36]))
var.cor <- 
  M.tbl %>%
  filter((abs(correlation) > 0.5) & (abs(correlation) < 1)) %>%
  arrange(desc(correlation))

ggplot(data = var.cor) +
  geom_bar(mapping = aes(x = reorder(variable, correlation),
                         y = correlation),
           stat = "identity",
           fill = "#27ae60",
           color = "#16a085",
           width = 0.5) +
  labs(x = "Variable",
       y = "Correlation") +
  coord_flip()
```

## Linear model

```{r}
lm.model <- lm(SalePrice ~ ., 
               data = train)
lm.model.summary <- summary(lm.model)
lm.model.summary

c <- coef(lm.model.summary)
c <- 
  as_tibble(data.frame(var = rownames(c), c, row.names = NULL)) %>%
  filter(Pr...t.. < 0.01) %>%
  arrange(Pr...t..)
c
```

## Next up

* Stepwise regression
* Lasso regressin
* PCA
* More EDA on categorical and numerical variables
* Random Forest
* XgBoost

1. Note: Ensure that the sign of the slope parameter is correct

2. Table with sign p-value for every column selected
Check for collinearity because of sign

3. Evaluation metrics - RMSE, 
